Single Site, Single Disturbance, MultiPool

### 2020-11-17 SingleSite_SingleDisturb_MultiPool.Rmd
SSSD-MV: added disturbed EnKF and ensemble adjustment to both EnKFs

```{r}
## spin up
ne = 100
met = sample.int(nd,ne,replace=TRUE)
par = sample.int(np,ne,replace=TRUE)
Na <- matrix(NA,ne,3)
for(i in 1:ne){
  Na[i,] = VSEM(as.numeric(Theta[par[i],]),PAR = PAR[met[i],])[nt,2:4]
}
Na.orig = Na
Nmu = Nf = Na*NA
par(mfrow=c(1,3))
hist(Na[,1],main="Cleaf")
hist(Na[,2],main="Csoil")
hist(Na[,3],main="Cstem")
```

@@

## storage
Dbar = 0
Analysis <- Forecast <- list()
removal = matrix(0,ne,NT)

@@

for(t in 1:NT){

@@

  ## Disturbance
  d.prior <- rbinom(ne,1,p.dist)             ## prior on disturbance status
  sel <- !as.logical(d.prior)                ## select undisturbed ensemble members
  if(sum(!sel)>2){
    sel.dist = which(!sel)
    for(j in sel.dist){                      ## run disturbance model
      Nf[j,] = disturbance(Nf[j,])
    }
    mufD = colMeans(Nf[!sel,])
    pfD = solve(diag(c(0.15,0.5,0.25)^2)) #try(solve(var(Nf[!sel,])))
  } 
  if(sum(!sel) <= 2 || class(pfD) == "try-error") {   ## backup values if the ensemble stats fail
      mufD = c(mu0[1],mean(Nf[,2]),mu0[2])             
      pfD  = solve(diag(c(0.15,0.5,0.25)^2))
  }
  Nf <- tobit(Nf)

@@

  ## draw new IC     *** REPLACE WITH ENSEMBLE ADJUSTMENT ***
  sel.ic <- sample.int(nrow(dat), ne, replace=TRUE)
  Na <- dat[sel.ic,nsel]
}

@@

  muf = colMeans(Nf[sel,])    ## Not disturbed forecast means
  Pf  = var(Nf[sel,])
  pf  = solve(Pf)             ## Not disturbed forecast precision
  Forecast.EnKF[[t]] <- list(d.prior=d.prior,Nf=Nf,muf=muf,pf=pf)

@@

  ## draw new IC     *** REPLACE WITH ENSEMBLE ADJUSTMENT ***
  Na <- mvtnorm::rmvnorm(ne,mua,Pa)

>>>>

```{r}
## spin up
ne = 100
met = sample.int(nd,ne,replace=TRUE)
par = sample.int(np,ne,replace=TRUE)
Na <- matrix(NA,ne,3)
for(i in 1:ne){
  Na[i,] = VSEM(as.numeric(Theta[par[i],]),PAR = PAR[met[i],])[nt,2:4]
}
Na.orig = Na
Nmu = NmuK = NfK = Nf = Na*NA
par(mfrow=c(1,3))
hist(Na[,1],main="Cleaf")
hist(Na[,2],main="Csoil")
hist(Na[,3],main="Cstem")
```

@@

## storage
Dbar = 0
Analysis <- Forecast <- list()
Forecast.KD = Analysis.KD = list() ## EnKF with Disturbance
removal = matrix(0,ne,NT)

@@

Na <- NaK <- Na.orig
THETAK <- THETA
for(t in 1:NT){

@@

  ## update IC
  THETA[,9:11]  <- Na    ## Multinomial
  THETAK[,9:11] <- NaK  ## EnKF w/ disturbance

  ## Forecast 
  for(i in 1:ne){
    Nmu[i,]  = VSEM(THETA[i,], PAR = PAR[met[i],tsel])[365,2:4] ## deterministic simulation
    NmuK[i,] = VSEM(THETAK[i,],PAR = PAR[met[i],tsel])[365,2:4] ## deterministic simulation
    epsilon  = mvtnorm::rmvnorm(1,rep(0,3),Q)                   ## process error  ** shared between versions **
    Nf[i,]   = Nmu[i,]  + epsilon
    NfK[i,]  = NmuK[i,] + epsilon
  }

  ## Disturbance
  d.prior <- rbinom(ne,1,p.dist)             ## prior on disturbance status
  sel <- !as.logical(d.prior)                ## select undisturbed ensemble members
  if(sum(!sel)>2){
    sel.dist = which(!sel)
    for(j in sel.dist){                      ## run disturbance model
      Nf[j,]  = disturbance(Nf[j,])
      NfK[j,] = disturbance(NfK[j,])
    }
    mufD  = colMeans(Nf[!sel,])
    pfD = solve(diag(c(0.15,0.5,0.25)^2)) #try(solve(var(Nf[!sel,])))
  } 
  if(sum(!sel) <= 2 || class(pfD) == "try-error") {   ## backup values if the ensemble stats fail
      mufD = c(mu0[1], mean(Nf[,2]),  mu0[2])            
      pfD  = solve(diag(c(0.15,0.5,0.25)^2))
  }
  Nf <- tobit(Nf)

  ## EnKF forecast stats
  NfK <- tobit(NfK)
  mufK = colMeans(NfK)
  PfK  = var(NfK)
  pfK  = solve(PfK)             ## Not disturbed forecast precision
  Forecast.KD[[t]] <- list(d.prior=d.prior,Nf=NfK,muf=mufK,pf=pfK)

@@

  ## draw new IC     *** REPLACE WITH ENSEMBLE ADJUSTMENT ***
  sel.ic <- sample.int(nrow(dat), ne, replace=TRUE)
  Na <- dat[sel.ic,nsel]

  ## EnKF Analysis
  I = H <- diag(rep(1,3))
  K <- PfK %*% t(H) %*% solve(H%*%PfK%*%t(H) + R)
  muaK <- mufK + K%*%(No[t,] - mufK) 
  PaK <- (I-K %*% H)%*%PfK    
  paK = solve(PaK)
  NaK <- PEcAn.assim.sequential::adj.ens(Pf = PfK,X = NfK,mu.f = mufK,mu.a = muaK,Pa = PaK)
  Analysis.KD[[t]] <- list(mua=muaK,pa=paK,Pa=PaK,Na=NaK)

@@

  muf = colMeans(Nf)    ## Not disturbed forecast means
  Pf  = var(Nf)
  pf  = solve(Pf)             ## Not disturbed forecast precision
  Forecast.EnKF[[t]] <- list(d.prior=d.prior,Nf=Nf,muf=muf,pf=pf)

@@

  ## draw new IC     *** REPLACED WITH ENSEMBLE ADJUSTMENT ***
#  Na <- mvtnorm::rmvnorm(ne,mua,Pa)
  Na <- PEcAn.assim.sequential::adj.ens(Pf = Pf,X = Nf,mu.f = muf,mu.a = mua,Pa = Pa)
